---
title: "Narrative Cloze Task: Experiment 3"
author: |
  | Erin Bennett
  | `erindb@stanford.edu`
bibliography: ../bibliography.bib
output:
  pdf_document:
    toc: false
    highlight: zenburn
    toc_depth: 3
---

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, sanitiz =TRUE, fig.width=5, fig.height=3)
```

```{r}
library(tidyr)
library(dplyr)
library(pwr)
library(rjson)
library(ggplot2)
library(ggthemes)
library(entropy)
theme_new <- theme_set(theme_few())

# for bootstrapping 95% confidence intervals
theta <- function(x,xdata) {mean(xdata[x])}
ci.low <- function(x) {
  quantile(bootstrap::bootstrap(1:length(x),1000,theta,x)$thetastar,.025)}
ci.high <- function(x) {
  quantile(bootstrap::bootstrap(1:length(x),1000,theta,x)$thetastar,.975)}

options(digits=3)
```

# Experiment

The narrative cloze task is a common evaluation task for script induction. I would like to understand humans' performance on the narrative cloze task.

In this task, participants read stories involving a particular entity. One of the events the entity is involved in is left blank. Participants are asked to fill in this blank with a plausible completion.

```{r}
# raw.data.path = "../data-collection/experiment3/production-results/"
# all.data <- data.frame()
# files <- dir(raw.data.path,pattern="*.json")
# 
# for (file.name in files) {
#   json_file <- readLines(paste(raw.data.path,file.name,sep=""))
#   json_file_str = paste(json_file, collapse = "")
#   json_file_str = gsub(",}", "}", json_file_str)
#   jso = jsonlite::fromJSON(json_file_str)
#   
#   # events = ((jso$answers)$data)$events
#   # minutes = (max(events$time) - min(events$time)) / 60000
#   
#   trials = ((jso$answers)$data)$trials
#   demographics = ((jso$answers)$data)$demographics
#   for (i in 1:(nrow(demographics))) {
#     trials[demographics[i,'qtype']] = demographics[i, 'response']
#   }
#   jso1 <- data.frame(trials)
#   # jso1$total_minutes = minutes
#   jso1$workerid = strsplit(file.name, '.json')[[1]]
# 
#   all.data <- rbind(all.data, jso1)
# }
# 
# all.data = all.data %>% mutate(workerid = factor(as.numeric(factor(workerid))))
#                                # minutes = as.numeric(minutes))
# 
# # all.data %>% group_by(workerid) %>%
# #   summarise(condition = condition[1],
# #             minutes=total_minutes[1]) %>%
# #   as.data.frame %>%
# #   filter(minutes > 10) %>%
# #   mutate(bonus = round((minutes - 5)*0.05, 2))


```

<!--
## Demographics and attention checks


Comments:

```{r}
print(unique(all.data$comments))
```

Participants thought the study was about:


```{r}
print(unique(all.data$studyQuestionGuess))
```


Average completion time was `r round(mean(completion.time$minutes))` minutes.


```{r}
attention.checks = all.data %>% group_by(workerid) %>% summarise(attention.check = response[chain=='attention']=='apples') %>% as.data.frame
good.Ss = attention.checks$workerid[attention.checks$attention.check]
```

`r length(unique(all.data$workerid)) - length(good.Ss)` participants were excluded for failing an attention check.

-->

```{r}
# d = all.data %>% mutate(subid = workerid)
# condition.design = d %>% group_by(condition) %>% summarise(n = length(unique(subid)))
# print(condition.design)
# # d = d %>% select(workerid, document, chain, clozeIndex, original, gloss, response,
# #                  condition, trialnum)
# # write.csv(d, 'experiment3.csv', row.names=F)
```

```{r}
d = read.csv('experiment3.csv')
```

There were between `r min(condition.design$n)` and `r max(condition.design$n)` participants in each of 3 conditions: full text for all entity-relevant sentences, "caveman" speak for all entity-relevant sentences, and "event" only sentences with only entity-role and main verb.

# Results

```{r}
```

# References